Crowdsourced geometric morphometrics enable rapid large-scale collection and

### 2015-09-16 manuscript/manuscript.Rmd
Code drop -- preprint version

---
title: Crowdsourced geometric morphometrics enable rapid large-scale collection and
  analysis of phenotypic data
author:
- affiliation: ucla
  email: jonathan.chang\@ucla.edu
  footnote: Author for correspondence jonathan.chang\@ucla.edu
  name: Jonathan Chang
- affiliation: ucla
  name: Michael E. Alfaro
bibliography:
- library.bib
- knitcitations.bib
output:
  pdf_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 7
    keep_tex: yes
    latex_engine: xelatex
    template: inc/mstemplate.tex
  html_document:
    toc: yes
  word_document: default
documentclass: inc/msarticle
fontsize: 11pt
geometry: margin=1.2in
layout: preprint
lineno: yes
linestretch: 1.05
csl: inc/methods-in-ecology-and-evolution.csl
running_head: Fast crowdsourced phenotypic data collection
address:
- address: |
    Department of Ecology and Evolutionary Biology, University of California, Los Angeles, CA, USA
  code: ucla
word_count: 6000?
---

```{r prelude, echo=FALSE, message=FALSE, warning=FALSE}
library("MASS")
library("readr")
library("dplyr")
library("ggplot2")
library("stringr")
library("magrittr")
library("tidyr")
library("reshape2")
library("scales")
library("geomorph")
library("rjson")
library("ipred")
library("vegan")
library("lubridate")
library("directlabels")
library("BAMMtools")
library("knitr")
library("kfigr")
library("knitcitations")
library("fish.reliability")
options("citation_format"="pandoc")
cleanbib()
set.seed(111) # reproducible and stable sampling
data(fish_reliability)
data(fish_families)
knitr::opts_chunk$set(echo=FALSE, cache=1, fig.align="center")
```

@@

3. We find that crowdsourced workers perform similarly to experienced morphologists on the same digitization tasks. We also demonstrate the speed and accuracy of our method on 7 families of ray-finned fishes (Actinopterygii), and find that rates of body shape evolution are correlated with diversification rates.

4. Crowdsourcing will enable the collection of morphological data across vast radiations of organisms, and will facilitate richer inference on the macroevolutionary processes that shape phenotypic diversity across the tree of life.

@@

Integrating phenotypic data, such as anatomy, behavior, physiology, and other traits, with phylogenies is powerful workflow for investigating the patterns of biological evolution. Recent advances in next-generation sequencing [@Shendure2008; @Meyer2008] and sequence capture technologies [@Faircloth2012b; @Lemmon2012] have made phylogenetic inference of large radiations of organisms possible [@McCormack2012; @McCormack2013; @Faircloth2013; @Faircloth2014].
These large phylogenies have been successfully used in conjunction with existing comprehensive phenotypic datasets to answer key questions about the mode and tempo of diversification and morphological evolution [@Rabosky2013]. However, similar breakthroughs for generating new phenotypic datasets have been comparatively uncommon, likely due to the high expense and effort required [reviewed in @Burleigh2013].

Creating these large phenotypic datasets has generally required an extended dedicated effort of measuring and describing morphological or behavioral traits that are then coded into a comprehensive data matrix. One such example is the Phenoscaping project [<http://kb.phenoscape.org>; @Deans2015], and its complementary efforts in the Vertebrate Taxonomy Ontogeny [@Midford2013] and Hymenoptera Anatomy Ontology [@Yoder2010], which requires large amounts of researcher effort to compile. Other approaches include using machine learning [@Mckinney2013a] or natural language processing [@Cui2012a] to identify or infer phenotypes. These statistical techniques function ideally with either a large training dataset or a complex model [@Brill2003; @Halevy2009; @Hastie2009], both of which also require intensive researcher effort to build and validate. Finally, methods such as high-throughput infrared imaging, mass spectrometry, and chromatography have been successfully used in plant physiology [@Furbank2011]
 and microbiology [@Skelly2013], but these methods may not be applicable for zoological researchers.
These approaches all share a similar goal of collecting large comparative datasets, but also require large investments in researcher effort. This bottleneck has limited the taxonomic breadth and depth of research in comparative biology.

Here we present a method and toolkit to efficiently collect two-dimensional geometric morphometric phenotypic data at a "phylogenomic" scale. 
We developed a novel web browser-based image landmarking application, and use Amazon Mechanical Turk (<https://www.mturk.com>) to distribute digitization tasks to remote workers (hereafter *turkers*) over the Internet, who are paid for their contributions.
We evaluate the accuracy and precision of turkers by assigning identical image sets and digitization protocols to users who are experienced with fish morphology (hereafter *experts*), and compare the inter- and intra-observer differences between turkers and experts.
As an example of how this new method may be used, we construct a phylogenetic analysis pipeline to download photographs and phylogenies of 7 actinopterygiian families from the web, collect Mechanical Turk shape results, and analyze the rate of diversification and body shape evolution using BAMM [@Rabosky2014a].

@@

Amazon Mechanical Turk ("MTurk") is a web-based service where Requesters can request work, known as Human Intelligence Tasks ("HITs") to be performed by Workers. Workers work from home and submit the tasks over the Internet, where Requesters review it, and, if they are satisfied with the results, accept the work and pay the Worker. We use MTurk as a platform to distribute our geometric morphometric tasks and financially compensate the worker accordingly. Scientific collection of data over MTurk and similar services has generally been limited to the fields of psychology and computer science, and there have been few attempts to crowdsource biological trait data [but see @Burleigh2013].

@@

We developed an application that runs completely on the user's local web browser. This simplifies the infrastructure challenge of needing to serve many crowdsdourced workers simultaneously. The web application is configured with a simple JavaScript Object Notation (JSON) file that describes the landmarks necessary to complete an image digitization task (for more details, see Supplementary Material).

Although digitizing and landmarking a single image [microtasks *sensu* @Good2013] is most effective for high-throughput work on MTurk, it is unsuitable for conducting controlled experiments. To solve this issue we created a server-side application backend that automatically distribute tasks according to a configurable set of images and experimental protocol. This application also mimics an official Amazon Mechanical Turk endpoint, to decouple the experimental backend from the image landmarking frontend. External non-MTurk workers can also participate in the same experiment, ensuring effective comparisons across these groups. The software is open source and uses PHP and SQLite, and can be installed on nearly all web hosting services (<https://github.com/jonchang/fake-mechanical-turk>).

@@

Collecting landmark-based geometric morphometric data at scale permits detailed analysis of different sources of error, such as among- and within-observer variation [@VonCramon-Taubadel2007]. To assess whether the quality of data gathered by workers recruited through Amazon Mechanical Turk was significantly different than traditionally-collected data, we asked both turkers and experts to landmark a set of 5 fish images, 5 times each. All participants used the same protocol and same software to digitize the same set of fishes.

The landmarks were chosen according to previously-published literature concerning fish shape [@Fink1995; @Cavalcanti1999; @Ruber2001a; @Klingenberg2003; @Chakrabarty2005; @Frederich2008; @Thacker2014], and included landmarks both easy to digitize, such as the anterior margin of the eye, and challenging to digitize, such as the opercular series. We used 5 replicate measurements per worker per image to quantify measurement error within and among observers [@Palmer1986] using *geomorph* `r citep(citation("geomorph"))`.

We used the median shape of the combined experts' data as the "true" shape for all comparisons. To assess the accuracy of worker data, as compared

To assess the differences between turker and expert data, we compared the median turker position to the median expert position of each landmark. We assumed that the expert median was indeed the true position of that landmark, and calculated the absolute Euclidian distance. Larger distances would indicate low turker accuracy, while smaller distances would indicate high turker accuracy. We also examined the variance in turker landmarks. For each landmark, we rotated the cluster of points and took calculated the median absolute deviation between points along a single dimension. This

To assess within-individual variation, we compared each worker's shape replicate to their own median shape. To assess within-group and between-group variation, we also compare those individual worker median shapes to the median shape of all other turkers and all other experts. We also used the median absolute deviation (MAD) as a dispersion statistic for within-worker, within-group, and between-group comparisons. This is a statistic robust to outliers

The curse of dimensionality is a common problem in machine learning, where $p$, the number of predictors, greatly exceeds $n$, the number of observations. Geometric morphometric datasets are typically extremely highly dimensional, with $2 * k - 1$ observations per 2D landmark. To reduce the dimensionality of our dataset, we orthogonalize our features used a generalized Procrustes superimposition as implement in the R package geomorph. We then use logistic regression on the PC axes, starting with PC1 and adding each subsequent axis in a stepwise fashion. We validate each model using $K$-fold cross-validation, which splits the data into $K$ equally sized groups, using $K-1$ groups to train the model and the last group to validate the model. This process is repeated for each subset of the groups. We set $K=10$ to validate the model, for its balanced statistical properties [@Kohavi1995; @Hastie2009]. Using 10-fold CV, we calculated the confusion matrix for each model in turn, and plot the sensitivity vs specificify (also known as the receiver operator characteristic curve).

To reduce the effect of the curse of dimensionality ($p >> n$, where the number of predictors  exceeds the number of observations), we first

<!-- (XXX: Talk about setting the consensus morphologist shape as the "true" shape in terms of determining accuracy / precision of individual workers and the consensus turker shape) -->

To determine whether turkers and experts were statistically distinguishable, we used the machine learning technique linear discriminant analysis [LDA, @Ripley1996] as implemented in *MASS* `r citep(citation("MASS"))`. A common error in assessing the efficacy of machine learning methods is training the model and then testing the model using the same data set for both. The researcher can make the model increasingly sophisticated to better fit the training dataset, but be unable to make accurate inferences to real-world data, due to overfitting. Conversely, withholding too much data for validation results in underfitting and the inability of the model to generalize its predictions to real data. To guard against these potentially misleading inferences, we used $K$-fold cross-validation (CV), which splits our data into $K$ equally-sized groups, using $K-1$ for training and $1$ for validation, and tests the model for all possible permutations of these groups. We used a value of $K=10$ to validate the LDA model, for its balanced approach to the bias-variance tradeoff [@Kohavi1995; @Hastie2009]. Using 10-fold CV, we assessed the LDA model's misprediction rate. Acceptable misprediction rates vary depending upon the application, here we judge 20% misprediction rate to sufficently distinguish turkers and experts, while a 50% or higher rate is no better than a random guess. We also used quadratic discriminant analysis (QDA), to test the robustness of our results with a different model. All analyses were performed in R `r citep(citation())`. (XXX: How to interpret the result of this method. Use R package caret for easier comparisons among different machine learning techniques. Try out support vector machines.)

We calculated the per-digitizer median shape for each image used, as well as the consensus turker and morphologist landmarks, and projected these shapes into Procrustes space. We calculated the 95% confidence ellipse to estimate the uncertainty around these shape estimates. (XXX: How to interpret the result of this method)

@@

A common strategy in fish comparative studies is to examine evolutionary dynamics within a single family (XXX: cite more papers that do this) [@Frederich2013; @Thacker2014], potentially due to the extensive amount of time necessary to collect data.

To demonstrate the utility of obtaining comparative data using this method, we downloaded previously-published phylogenies for 7 fish families (Acanthurdiae, Apogonidae, Balistidae, Chaetodontidae, Labridae, Pomacentridae, and Tetraodontidae) from TreeBASE and Dryad (XXX: CITE TREES -- need to get info from Mike) and matched them to left-lateral images from the Encyclopedia of Life (<http://eol.org/>) using the EOL API [@Parr2014]. If there were multiple images for a single species, we manually selected the best one. Crowdsourced workers collected body shape data following a standard protocol (Supplementary Material). We then determined the major components of shape variation using a Procrustes-aligned principal components analysis (PCA) [@Mardia1979; @Bookstein1991] as implemented in the R package *geomorph* `r citep(citation("geomorph"))`.

To test whether our method can improve on the case where the data collection method is geometric morphometrics, we use the average time it took an expert to measure a single fish image and predict the time it would take for a single individual expert to measure our 7 fish families with 5 replicates, and compare it to the time it actually took turkers to collect these measurements at the same replication level. If the turkers in aggregate annotated images more quickly than a single expert would have, this suggests that the parallelization afforded by crowdsourcing is effective at reducing the time required for data collection.

We used Bayesian Analysis of Macroevolutionary Mixtures [*BAMM*, @Rabosky2014a] to estimate the diversification rate as well as the rate of morphological evolution, retaining the PC axes whose eigenvalues exceeded the corresponding random broken-stick component [@Jackson1993; @Legendre1998], for each of the 7 families. BAMM estimates the location of rate shifts in either diversification or character evolution using a transdimensional (reversible jump) Markov Chain Monte Carlo method, where the number of rate regimes on a phylogeny is allowed to vary. We assessed convergence and mixing using Tracer [@rambaut2007tracer]. We then used BAMMtools `r citep(citation("BAMMtools"))` to detect whether body shape affects change in the speciation or extinction rate. State-dependent diversification has often been invoked  to explain the high species richness enjoyed by certain lineages [@Mitter1988; @Barraclough1998]; this pipeline demonstrates a largely-automated way to detect these. `r citep(bibentry(bibtype="Article", title="A robust semi-parametric test for detecting trait-dependent diversification", author=c("Daniel L. Rabosky", "Huateng Huang"), year="in.review", journal="Systematic Biology"))`

@@

```{r turker_v_expert, fig.cap="Test figure caption", anchor="figure"}
no_st <- fish_reliability %>% filter(role != "student") %>% mutate(role = factor(role))

@@

all_stats <- left_join(accuracy, precision) %>% left_join(full_family) %>% mutate(outlier_label = ifelse(outlier, as.character(mark), NA))

ggplot(all_stats, aes(median, mad, color=outlier)) + geom_point() + labs(x="Accuracy", y="Precision: log2 turker variance / expert variance ") + facet_wrap(~full_family)# + geom_dl(aes(label=outlier_label), size=1, list("top.bumptwice", dl.trans(y=y+0.1), cex=0.75)) + scale_color_manual(values=c("black", "red")) + ylim(-max(abs(all_stats$mad) + 0.25), max(abs(all_stats$mad) + 0.25)) + theme("legend.position" = "none")
```

```{r geomorph_lm, eval=FALSE}
fmted <- format_for_gpagen(no_st)
allsplit <- str_split_fixed(dimnames(fmted)$id, "_", 3)
family_vec <- allsplit[, 1]
role_vec <- allsplit[, 2]
ind_vec <- allsplit[, 3]
gpaed <- gpagen(fmted, ShowPlot=F)

role_by_ind <- procD.lm(gpaed$coords ~ family_vec * role_vec * ind_vec, iter=999, RRPP=T)
ind_only <- procD.lm(gpaed$coords ~ family_vec:ind_vec, iter=999, RRPP=T)

advanced.procD.lm(gpaed$coords ~ ind_vec,  ~ role_vec * ind_vec, ~ family_vec, iter=499)
```

```{r median_tbl, eval=F, results="asis", anchor="table", fig.cap="Median table"}
med_wide <- meds %>% spread(key=family, value=value)
med_wide %>% knitr::kable(digits = 2, caption = "Median table")
```

```{r mad_tbl, eval=F, results="asis", anchor="table", fig.cap="MAD table"}
mad_wide <- mads %>% spread(key=family, value=value)
mad_wide %>% knitr::kable(digits = 2, caption="MAD table")
```

We compare the difference in accuracy and precision between turkers and experts in Figure `r figr("turker_v_expert")`.

The most accurate and precise points are those that are related to the position of the eye (landmarks E1 and E2). The least accurate are those in the opercular series (O1-O5), particularly the ones related to the preopercle (O1-O3) as well as in species where the preopercle is difficult to visualize from external morphology alone (e.g., Tetraodontidae). We also plot the median difference and the

```{r cv_table, results="asis", anchor="table", message=FALSE}

subsamp <- group_by(no_st, family, role) %>% sample_n(size=400) %>% ungroup()

>>>>

---
title: Crowdsourced geometric morphometrics enable rapid large-scale collection and
  analysis of phenotypic data
author:
- affiliation: ucla
  email: jonathan.chang\@ucla.edu
  footnote: Author for correspondence jonathan.chang\@ucla.edu
  name: Jonathan Chang
- affiliation: ucla
  name: Michael E. Alfaro
bibliography:
- library.bib
- knitcitations.bib
output:
  pdf_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 7
    keep_tex: yes
    latex_engine: xelatex
    template: inc/mstemplate.tex
  html_document:
    toc: yes
  word_document: default
documentclass: inc/msarticle
endfloat: yes
fontsize: 11pt
geometry: margin=1.2in
layout: preprint
lineno: yes
linestretch: 2
csl: inc/methods-in-ecology-and-evolution.csl
running_head: Fast crowdsourced phenotypic data collection
address:
- address: |
    Department of Ecology and Evolutionary Biology, University of California, Los Angeles, CA, USA
  code: ucla
word_count: 7000
---

```{r prelude, echo=FALSE, message=FALSE, warning=FALSE}
library("MASS")
library("readr")
library("dplyr")
library("ggplot2")
library("stringr")
library("magrittr")
library("tidyr")
library("reshape2")
library("scales")
library("geomorph")
library("rjson")
library("ipred")
library("vegan")
library("lubridate")
library("directlabels")
library("BAMMtools")
library("knitr")
library("kfigr")
library("knitcitations")
library("parallel")
library("fish.reliability")
options("citation_format"="pandoc")
cleanbib()
set.seed(111) # reproducible and stable sampling
data(fish_reliability)
data(fish_families)
knitr::opts_chunk$set(echo=FALSE, cache=1, fig.align="center")
```

@@

3. We find that crowdsourced workers perform similarly to experienced morphologists on the same digitization tasks. We also demonstrate the speed and accuracy of our method on seven families of ray-finned fishes (Actinopterygii).

4. Crowdsourcing will enable the collection of morphological data across vast radiations of organisms, and can facilitate richer inference on the macroevolutionary processes that shape phenotypic diversity across the tree of life.

**Keywords**: crowdsourcing, morphometrics, phenotyping, morphology, comparative methods, macroevolution, Actinopterygii

@@

Integrating phenotypic data, such as anatomy, behavior, physiology, and other traits, with phylogenies is powerful strategy for investigating the patterns of biological evolution. Recent advances in next-generation sequencing [@Shendure2008; @Meyer2008] and sequence capture technologies [@Faircloth2012b; @Lemmon2012] have made phylogenetic inference of large radiations of organisms possible [@McCormack2012; @McCormack2013; @Faircloth2013; @Faircloth2014]. However, similar breakthroughs for generating new phenotypic datasets have been comparatively uncommon, likely due to the high expense and effort required [reviewed in @Burleigh2013].

Creating these large phenotypic datasets has generally required an extended dedicated effort of measuring and describing morphological or behavioral traits that are then coded into a comprehensive data matrix. One such example is the Phenoscaping project [<http://kb.phenoscape.org>; @Deans2015], and related efforts in the Vertebrate Taxonomy Ontogeny [@Midford2013] and Hymenoptera Anatomy Ontology [@Yoder2010], which require large amounts of researcher effort to collate. Other approaches include using machine learning [@Dececchi2015], machine vision [@Corney2012; @Corney2012a], or natural language processing [@Cui2012a] to identify or infer phenotypes. These statistical techniques function ideally with either a large training dataset (e.g., a predefined ontogeny database) or a complex model [@Brill2003; @Halevy2009; @Hastie2009], both of which also require intensive researcher effort to build and validate. Finally, methods such as high-throughput infrared imaging, mass spectrometry, and chromatography have been successfully used in plant physiology [@Furbank2011]
 and microbiology [@Skelly2013], but these methods may not be applicable for zoological researchers.
These approaches all share a similar goal of collecting large comparative datasets, but also require large investments in researcher effort. This bottleneck in researcher availability has limited the scope of work in comparative biology.

Although it is now possible to build phylogenetic trees with thousands of tips, and phenotypic data sets have similarly been growing larger and larger, the traits that are typically studied at this scale tend to be simple: geographic occurrences [@Jetz2012], one or two continuous characters [@Harmon2010; @Rabosky2013], a single discrete character [@Goldberg2010; @Edwards2011; @Price2012], or some combination of these [@Zanne2014; @Pyron2014]. A richer understanding of the forces that shape macroevolution requires the collection of more detailed phenotypic trait data at scale.

Here we present a method and toolkit to efficiently collect two-dimensional geometric morphometric phenotypic data at a high-throughput "phenomic" scale. 
We developed a novel web browser-based image landmarking application, and use Amazon Mechanical Turk (<https://www.mturk.com>) to distribute digitization tasks to remote workers (hereafter *turkers*) over the Internet, who are paid for their contributions.
We evaluate the accuracy and precision of turkers by assigning identical image sets and digitization protocols to users who are experienced with fish morphology (hereafter *experts*), and compare the inter- and intra-observer differences between turkers and experts.
To illustrate the efficiency of this approach, we construct a phylogenetic analysis pipeline to download photographs and phylogenies of seven actinopterygiian families from the web, collect Mechanical Turk shape results, analyze the rate of diversification and body shape evolution using BAMM [@Rabosky2014a], and compare the time required for this workflow to traditional approaches. We also discuss the role that crowdsourcing is best suited in large-scale morphological analyses, and suggest ways to integrate crowdsourced data as part of larger initiatives to digitize biodiversity.

@@

Amazon Mechanical Turk ("MTurk") is a web-based service where Requesters can request work, known as Human Intelligence Tasks ("HITs") to be performed by Workers. Workers work from home and submit the tasks over the Internet, where Requesters review it, and, if they are satisfied with the results, accept the work and pay the Worker. We use MTurk as a platform to distribute our geometric morphometric tasks and financially compensate the worker accordingly. Scientific collection of data over MTurk and similar services has generally been limited to the fields of psychology and computer science, and there have been few attempts to crowdsource biological trait data [@Burleigh2013].

@@

We developed an geometric morphometric digitization application that runs completely on the user's local web browser, using the HTML5 Canvas interface. This simplifies the infrastructure challenge of needing to serve many crowdsourced workers simultaneously, since workers will not need to download desktop software such as tpsDig (<http://life.bio.sunysb.edu/ee/rohlf/software.html>) before generating data. The web application is configured with a simple JavaScript Object Notation (JSON) file that describes the landmarks necessary to complete an image digitization task (Supplemental Figure S1). Point landmarks, semilandmark curves, and linear measurements are all supported. The software is available at <https://github.com/jonchang/eol-mturk-landmark>.

Although digitizing and landmarking a single image [microtasks *sensu* @Good2013] is effective for high-throughput work on MTurk, it is unsuitable for conducting controlled experiments. To solve this issue we also created a server-side application backend that automatically distributes tasks according to a configurable set of images and experimental protocol. This application mimics an official Amazon Mechanical Turk interface endpoint, to facilitate drop-in replacement for an existing MTurk workflow. External non-MTurk workers can also participate in the same experiment, ensuring consistent comparisons across separate groups. The software is available at <https://github.com/jonchang/fake-mechanical-turk>.

@@

Collecting landmark-based geometric morphometric data at scale permits detailed analysis of different sources of error, such as among- and within-observer variation [@VonCramon-Taubadel2007]. To assess whether the quality of data gathered by workers recruited through Amazon Mechanical Turk was significantly different than traditionally-collected data, we asked turkers ($n = `r nrow(filter(fish_reliability, role == "turker") %>% distinct(sha1mac))`$) and experts ($n = `r nrow(filter(fish_reliability, role == "morphologist") %>% distinct(sha1mac))`$) to landmark a set of five fish images, five times each. All participants used the same protocol and same software to digitize the same set of fishes. The landmarks were carefully selected based on previously-published literature concerning fish shape [Supplemental Figure S2; @Fink1995; @Cavalcanti1999; @Ruber2001a; @Klingenberg2003; @Chakrabarty2005; @Frederich2008; @Claverie2014; @Thacker2014]. We also ensured that the chosen landmarks included morphological features that were relatively straightforward to digitize (the position of the eye) and features that were likely to be more challenging to digitize (the position of the preopercle bone), in order to test for turker and expert differences over a spectrum of difficulties. We report the inter-observer reliability for turkers and experts by computing the ratio of the among-individual and the sum of the among-individual and measurement error variance components in a repeated measures nested MANOVA [@Palmer1986; @Zelditch2012].

To assess the differences between turker and experts on a per-landmark basis, we first compared the median turker position to the median expert position of each landmark. We assumed that the expert median was the true position of that landmark, and calculated the absolute Euclidian distance. Larger distances would indicate low turker accuracy, while smaller distances would indicate high turker accuracy. We then examined the variance in turker landmarks. For each landmark, we rotated the cloud of points to maximize variance in one dimension, and calculated the log-ratio of median absolute deviations (MAD) between turkers and experts. This rotation is a conservative approach for assessing the difference in variance between these two groups, because it maximizes any apparent differences in landmark position. A positive log-ratio indicated that experts had lower variance than turkers, while a negative log-ratio indicated that turkers had lower variance. For all subsequent analysis, we excluded landmarks where turkers performed especially poorly, where either the accuracy or precision components for a given landmark exceeded 1.5 times the interquartile range of that component.

To determine whether turkers and experts were statistically distinguishable, we performed a non-parametric MANOVA using the randomized residual permutation procedure (RRPP) with 1,000 iterations [@Collyer2014]. The RRPP method reduces the effect of the "curse of dimensionality" ($p >> n$, where the number of predictors greatly exceeds the number of observations), a common problem in geometric morphometrics, and has been shown to have increased statistical power compared to a method where the raw data are randomized instead [@Anderson2003]. 
We test for a difference between mean turker and expert shapes against a null model of no difference between turker and expert changes, taking into account species-specific differences. 
A difference between models was considered significant if the p-value was less than $\alpha = 0.05$.

As a separate test, we use linear discriminant analysis [LDA, @Ripley1996], a statistical classification algorithm that finds features to differentiate between different classes of data, in this case turkers and experts. We assessed the accuracy of the LDA classification using 10-fold cross validation (CV), which splits our data into 10 equally-sized groups, using nine for training and one for validation [@Kohavi1995; @Hastie2009]. An acceptable misclassification rate varies depends on application, but here we use a 25% misprediction rate as a standard for sufficient accuracy. This is a highly forgiving standard, since a 50% misprediction rate is no better than a coin flip, and a 25% misprediction rate would still erroneously classify one in four turkers as experts or vice versa. We also use quadratic discriminant analysis (QDA), which relaxes some of the assumptions of LDA, and similarly report the QDA misclassification rate.

We calculated the per-individual median shape for each species used, as well as the consensus turker and morphologist shapes, and projected these shapes into Procrustes space, to visualize the orthogonalized differences in median shape among and between the types of digitizers.

@@

A common strategy in fish comparative studies is to examine evolutionary dynamics within a single family [@Ferry-Graham2001a; @Alfaro2005; @Alfaro2007; @Rocha2008; @Hernandez2009; @Dornburg2011; @Sorenson2013; @Santini2013; @Frederich2013; @Claverie2014; @Thacker2014], potentially due to the extensive amount of time necessary to collect data. 
To test whether our method can improve on the case where the data collection method is geometric morphometrics, we use the average time it took an expert to measure a single fish image and predict the time it would take for a single individual expert to measure all images at 5x replication, and compare it to the time it took turkers to collect these measurements at the same replication level. If the turkers in aggregate annotated images more quickly than a single expert would have, this suggests that the parallelization afforded by crowdsourcing is effective at reducing the total time required for data collection.

To demonstrate the utility of obtaining comparative data using this method, we use previously published phylogenies for seven fish families: Acanthuridae [@Sorenson2013], Balistoidae, Tetraodontidae [@Santini2013], Apogonidae, Chaetodontidae, Labridae [@Cowman2011; @Choat2012], and Pomacentridae [@Frederich2013]. We matched `r nrow(distinct(fish_families, tip))` species to left-lateral images from the Encyclopedia of Life (<http://eol.org/>) using their application programming interface [@Parr2014]. Crowdsourced workers placed landmarks describing body shape variation following a standard protocol (Supplementary Material). The Cartesian position of these landmarks were used in a generalized Procrustes analyses [@Gower1975; @Rohlf1990], which centers, scales, and rotates landmark configurations to minimize the least-squares distance between shapes. We then determined the major components of shape variation using a Procrustes-aligned principal components analysis (PCA) [@Mardia1979; @Bookstein1991] with the R package *geomorph* `r citep(citation("geomorph"))`, and used these principal components axes for subsequent analyses.

We used Bayesian Analysis of Macroevolutionary Mixtures [BAMM; @Rabosky2014a] to estimate rates of speciation and body shape evolution for all seven families. For the characters describing body shape, we use the PC axes whose eigenvalues exceeded the corresponding random broken-stick component [@Jackson1993; @Legendre1998]. BAMM estimates the location of rate shifts in either diversification or character evolution using a transdimensional (reversible jump) Markov Chain Monte Carlo method that samples a variety of models of lineage diversification and trait evolution. We assessed convergence and mixing using Tracer [@rambaut2007tracer]. We also repeated each analysis and simulated under the prior (without data) to exclude rate heterogeneity that occurred solely due to stochastic processes. We use a Bayes Factor criterion of $BF > 5$ to enumerate the set of credible shifts [@Shi2015] and visualized them in R using BAMMtools `r citep(citation("BAMMtools"))`.

@@

```{r geomorph_repeatability, warning=FALSE}
only_good_points <- fish_reliability %>% filter(role != "student", !mark %in% c("A3", "A4", "D3", "D4", "O1", "O2", "O3","O4","O5", "P3", "P4"))
reliability <- list()

res <- only_good_points %>% gpa_with_vectors(family, role, sha1mac, sequence)
vec <- res$vectors %>% group_by(family, sha1mac) %>% mutate(sequence = row_number(sequence))
full_model <- procD.lm(res$coords ~ vec$family + vec$family:vec$sha1mac:vec$sequence, iter = 1)
reliability$all <- ((full_model$MS[1] - full_model$MS[2])/5) / (full_model$MS[2] + ((full_model$MS[1] - full_model$MS[2])/5))

res <- only_good_points %>% filter(role == "turker") %>% gpa_with_vectors(family, role, sha1mac, sequence)
vec <- res$vectors %>% group_by(family, sha1mac) %>% mutate(sequence = row_number(sequence))
full_model <- procD.lm(res$coords ~ vec$family + vec$family:vec$sha1mac:vec$sequence, iter = 1)
reliability$turker <- ((full_model$MS[1] - full_model$MS[2])/5) / (full_model$MS[2] + ((full_model$MS[1] - full_model$MS[2])/5))

res <- only_good_points %>% filter(role == "morphologist") %>% gpa_with_vectors(family, role, sha1mac, sequence)
vec <- res$vectors %>% group_by(family, sha1mac) %>% mutate(sequence = row_number(sequence))
full_model <- procD.lm(res$coords ~ vec$family + vec$family:vec$sha1mac:vec$sequence, iter = 1)
reliability$expert <- ((full_model$MS[1] - full_model$MS[2])/5) / (full_model$MS[2] + ((full_model$MS[1] - full_model$MS[2])/5))

res <- fish_reliability %>% filter(role == "morphologist") %>% gpa_with_vectors(family, role, sha1mac, sequence)
vec <- res$vectors %>% group_by(family, sha1mac) %>% mutate(sequence = row_number(sequence))
full_model <- procD.lm(res$coords ~ vec$family + vec$family:vec$sha1mac:vec$sequence, iter = 1)
reliability$expert_all <- ((full_model$MS[1] - full_model$MS[2])/5) / (full_model$MS[2] + ((full_model$MS[1] - full_model$MS[2])/5))

res <- fish_reliability %>% filter(role == "turker") %>% gpa_with_vectors(family, role, sha1mac, sequence)
vec <- res$vectors %>% group_by(family, sha1mac) %>% mutate(sequence = row_number(sequence))
full_model <- procD.lm(res$coords ~ vec$family + vec$family:vec$sha1mac:vec$sequence, iter = 1)
reliability$turker_all <- ((full_model$MS[1] - full_model$MS[2])/5) / (full_model$MS[2] + ((full_model$MS[1] - full_model$MS[2])/5))
rm(vec, res, full_model)
```

```{r turker_v_expert, fig.cap="Per-family breakdown of accuracy vs. precision for each landmark. Accuracy is represented as the difference between the median turker location for that landmark and the median expert location, with the expert location assumed to be the true location. Precision is represented as the log-ratio of median absolute deviations between turkers and experts. More positive numbers indicate better expert precision, whereas more negative numbers indicate better turker precision. Points highlighted in red are those determined to be outliers (1.5 $\times$ IQR). See Supplemental Information for a labeled version of this figure.", anchor="figure", message=FALSE, warning=FALSE}
no_st <- fish_reliability %>% filter(role != "student") %>% mutate(role = factor(role))

@@

all_stats <- left_join(accuracy, precision) %>% mutate(outlier_label = ifelse(outlier, as.character(mark), NA))

ggplot(all_stats, aes(median, mad, color=outlier)) + geom_point() + labs(x="Accuracy: median turker - median expert", y="Precision: log turker variance / expert variance ") + facet_wrap(~family) + scale_color_manual(values = c("black", "red")) + theme_bw() + theme(legend.position="none") + xlim(0, 230)
ggsave("turker_precision.pdf", width=14, height=14, scale = 1/3)

# ggplot(highlighted, aes(median, mad)) + geom_point(color = "white") + labs(x="Accuracy: median turker - median expert", y="Precision: log turker variance ? expert variance ") + theme_black() + theme(legend.position = "none") + xlim(0, 130)
# ggsave("turker_precision_black.pdf", width=14, height=14, scale = 1/3)
# 
# 
# highlighted <- all_stats %>% mutate(should_color = family == "Tetraodontidae" & str_detect(mark, "O"))
# 
# ggplot(highlighted, aes(median, mad, color=should_color, label=mark)) + geom_point() + geom_text(data = filter(highlighted, should_color) %>% mutate(median = median + 4, mad = mad + 0.2)) + scale_color_manual(values = c("gray80", "red")) + labs(x="Accuracy: median turker - median expert", y="Precision: log turker variance ? expert variance ") + theme_black() + theme(legend.position = "none") + xlim(0, 130)
# 
# ggsave("turker_precision_black_puffers.pdf", width=14, height=14, scale = 1/3)
# 
# highlighted <- all_stats %>% mutate(should_color = family == "Chaetodontidae" & str_detect(mark, "[AD]3"))
# 
# ggplot(highlighted, aes(median, mad, color=should_color, label=mark)) + geom_point() + geom_text(data = filter(highlighted, should_color) %>% mutate(median = median + 4, mad = mad - 0.2)) + scale_color_manual(values = c("gray80", "red")) + labs(x="Accuracy: median turker - median expert", y="Precision: log turker variance ? expert variance ") + theme_black() + theme(legend.position = "none") + xlim(0, 130)
# 
# ggsave("turker_precision_black_chaets.pdf", width=14, height=14, scale = 1/3)
# 
# highlighted <- all_stats %>% mutate(should_color = family == "Balistidae" & str_detect(mark, "D[12]"))
# 
# ggplot(highlighted, aes(median, mad, color=should_color, label=mark)) + geom_point() + geom_text(data = filter(highlighted, should_color) %>% mutate(median = median + 4, mad = mad - 0.2)) + scale_color_manual(values = c("gray80", "red")) + labs(x="Accuracy: median turker - median expert", y="Precision: log turker variance ? expert variance ") + theme_black() + theme(legend.position = "none") + xlim(0, 130)
# 
# ggsave("turker_precision_black_triggers.pdf", width=14, height=14, scale = 1/3)
# 
# 
# highlighted <- all_stats %>% mutate(should_color = median > 10)
# 
# ggplot(highlighted, aes(median, mad, color=should_color, label=mark)) + geom_point() + geom_text(data = filter(highlighted, should_color) %>% mutate(median = median + 4, mad = mad - 0.2)) + scale_color_manual(values = c("gray80", "red")) + labs(x="Accuracy: median turker - median expert", y="Precision: log turker variance ? expert variance ") + theme_black() + theme(legend.position = "none") + xlim(0, 130)

# + geom_dl(aes(label=outlier_label), size=1, list("top.bumptwice", dl.trans(y=y+0.1), cex=0.75)) + scale_color_manual(values=c("black", "red")) + ylim(-max(abs(all_stats$mad) + 0.25), max(abs(all_stats$mad) + 0.25)) + theme("legend.position" = "none")
```
